---
title: "R Notebook"
output: html_notebook
---
```{r}
c(2, 5, 7) + 3 # work
c("2", "5", "7") + 3 #won't work
convert2numeric <- as.numeric(c("2", "5", "7")) 
convert2numeric + 3
c(2, 5, 7) + 3 # work
c("2", "5", "7") + 3 #won't work
```

```{r}
mylist <- list(2, 5, 7)
mylist[[1]] <- mylist[[1]]+3
mylist[[2]] <- mylist[[2]]+3
mylist[[3]] <- mylist[[3]]+3
print(mylist)
```

```{r}
unlist(mylist) + 3
```
convert list to atomic vector

```{r}
examples <- list()
examples$unlist$source1 <-
  list("A", "B", "C")
examples$unlist$result1 <- 
  unlist(examples$unlist$source1)
examples$unlist$result1 <- 
  unlist(
    list("A", "B", "C")
  )
print(examples$unlist$source1)
print(examples$unlist$result1)

class(examples$unlist$source1)
class(examples$unlist$result1)
```

```{r}
examples$unlist$source2 <- 
  list(
    list("A", "B", list("C")),
    list("D"),
    "E"
  ) # nested list
examples$unlist$result2 <- 
  unlist(
    examples$unlist$source2
  )
print(examples$unlist$source2)
print(examples$unlist$result2)
class(examples$unlist$source2)
class(examples$unlist$result2)
```

```{r}
examples$unlist$source3 <- list(5, 6, 7)
examples$unlist$result3 <-
  unlist(
    examples$unlist$source3
  )

class(examples$unlist$source3) # list class
class(examples$unlist$result3) # numeric class
```

```{r}
namedList <- list(name="John", spouse="Mary")

unlist(
  namedList
) -> unlist_namedList

unlist_namedList
```

#4.5(exercise)

```{r}
participations <- 
list('session 3',c("session 1", "session 2", "session 3"),'session 3',c("session 2", "session 1"),c("session 3", "session 1"),c("session 3", "session 2", "session 1"),'session 2','session 1',c("session 2", "session 1", "session 3"),c("session 3", "session 1", "session 2"))
```

```{r}
# sessions that 1st student attended
participations[[1]] 
# sessions that 2nd student attended
participations[[2]]
```

#4.3.2
```{r}
as.list(
  c("A", "B", "C")
) #the same as
list(
  "A", "B", "C"
) #NOt the same as 
list(
  c("A", "B", "C")
)

# element names will be preserved
as.list(
  c(name="John", spouse="Mary")
)
```

#4.3.3
```{r}
lgl <- c(TRUE, FALSE) #logical
num <- c(0.2, 3, 0) #numerical
```

```{r}
as.character(lgl) #the same as 
c("True", "False")
as.character(num)#the same as 
c("0.2", "3", "0")
```

```{r}
num + 3
as.character(num) + 3
```

```{r}
tookRcourse <- c(TRUE, FALSE, TRUE, TRUE)
class(tookRcourse)
as.numeric(tookRcourse)# True=1, False=0
howManyTookR <- sum(
  as.numeric(tookRcourse)
)
sum(
  c(1, 0, 1, 1)
)
1+0+1+1
print(howManyTookR)
howManyTookR <- sum(
  tookRcourse
  )
print(howManyTookR)
```

```{r}
which(toolRcourse)#only logical vector
```

```{r}
c("1", "5") +3
as.numeric(
  c("1", "5")
) + 3
```

```{r}
ming_guoDates <- c("1100822","1090723") # 民國100-8-22，民國109-7-23

# 民國1年 is 1911
westernDates <- 
  as.numeric(ming_guoDates) +
  19110000

westernDates 

lubridate::ymd(westernDates)
```

```{r}
#onnumeric vector
#tell in the number is not zero
print(num)
as.logical(num)
```

```{r}
# only person with a non-zero wage has a job
wage <- c(3000, 2000, 0, 1000)
haveJob <- as.logical(wage)
haveJob
```

```{r}
whoHasAJob <- which(haveJob)
whoHasAJob

# the wages of those who have a job
wage[whoHasAJob]
```

```{r}
wage[c(T, T, F, T)]
```

```{r}
wage <- c(3000, 2000, 0, 1000)
#want to pick out 3000 and 1000
wage[c(1, 4)]
wage[c(T, F, F, T)]
```

```{r}
wage[c(1, 4)]
wage[c(4, 1)]
```

```{r}
fct <- factor(
  c('參','貮','貮','貮','壹','貮','參','貮','參','參'),
  levels=c('壹','貮','參')
  )

dt <- lubridate::ymd_hms(
  c('2012-08-25 19:36:00','2018-01-06 10:44:00','2010-03-08 00:56:00')
)
```

```{r}
print(fct)
as.character(fct)

class(fct)
typeof(fct)
as.numeric(fct) # the positions in levels
```

```{r}
persians <- factor(
  c('۶','۹','۸','۹','۶','۸','۸','۵','۹','۹'),
  levels=c("۵", "۶", "۷", "۸", "۹")
  )

print(persians)
as.numeric(persians)
typeof(persians)

as.character(persians)
print(persians)
```

```{r}
print(dt)
as.character(dt)

typeof(dt)
as.numeric(dt) # how many seconds past 1970-01-01 00:00
```

```{r}
dt
dt + 30 # add 30 seconds
dt + 60*60 # add 1 hour
```

```{r}
x0 <- 1595950405 # the number of seconds
lubridate::as_datetime(x0)
```

```{r}
location_history <- jsonlite::fromJSON("https://www.dropbox.com/s/db2vt4w9u2w7onx/Location%20History.json?dl=1")
```

```{r}
location_history$locations$timestampMs/(10**3)
class(location_history$locations$timestampMs)
```

```{r}
timeStampMs_as.numeric <- 
  as.numeric(
    location_history$locations$timestampMs
  )
timeStamp_inSecondUnit <-
  timeStampMs_as.numeric/(10**3)
timeStamp_dateTimeClass <-
  lubridate::as_datetime(
    timeStamp_inSecondUnit
  )
```

```{r}
a=3
b=2
answer=a+b
```

```{r}
answer <- 
{
  a=3
  b=2
  a+b
}
```

```{r}
answer <- 
{
  a=3
  b=2
  a+b
}
```

```{r}
{
  a=3
  b=2
  a+b
} -> answer 
```

Layout the algorithm code
```{r}
task_result <- {
  # step 1: 
  
  # step 2: 
  
  # final step:
}
```

```{r}
location_history <- jsonlite::fromJSON("https://www.dropbox.com/s/db2vt4w9u2w7onx/Location%20History.json?dl=1")
location_history$locations$timestampMs
lubridate::as_datetime(location_history$locations$timestampMs)
class(location_history$locations$timestampMs)
timeStamp_dateTimeClass <- 
  {
    # Step 1: convert character to numeric
    step1 <- as.numeric(
      location_history$locations$timestampMs
    )
    # Step 2: given step 1 result, change millisecond to second unit
    step2 <- step1/1000
    # Step 3: given step 2 result, convert numeric seconds to date/time class
    lubridate::as_datetime(step2)
  }
```

```{r}
#cancel 2C 1A
chr <- c('C','C','A','A','B','B','B','B','C')
as.factor(chr)
table(chr)
table(as.factor(chr)) # class conversion
factor(
  chr, levles=c("B", "C", "A")
)
```

```{r}
{
  #do table count on chr
  table(chr)->tb_chr
  tb_chr
  #ask computer to figure out the level sequence
  sort(tb_chr, decreasing=T)->
sorted_tb_chr
  sorted_tb_chr
  #ask computer to give us names after sorting
  names(sorted_tb_chr)
}
```

```{r}
newLevels <- 
{
  # table check
  table(chr) ->tb_chr
  # sort in decreasing order
  sort(tb_chr, decreasing = T)-> 
    sorted_tb_chr
  #take out table item names
  names(sorted_tb_chr)
}
fct_chr <- factor(chr, levels=newLevels)
table(fct_chr)
```

```{r}
x |> f() |>g(y) |h(z)
#use x to do f, then take the result 
#to do g, then take the result 
#to do h 
```

```{r}
newLevels <- 
{
  # table check
  chr |>table() -> tb_chr
  table(chr) ->tb_chr
  # sort in decreasing order
  tb_chr |>sort(decreasing=T) ->
    sorted_tb_chr
  sort(tb_chr, decreasing = T)-> 
    sorted_tb_chr
  #take out table item names
  sorted_tb_chr |> names()
  names(sorted_tb_chr)
}
fct_chr <- factor(chr, levels=newLevels)
table(fct_chr)
```

```{r}
newLevels <- 
{
  # table check
  chr |>table()|>
  # sort in decreasing order
 sort(decreasing=T)|>
  #take out table item names
 names()
 
}
fct_chr <- factor(chr, levels=newLevels)
table(fct_chr)
```

```{r}
|> |> |> |> 
```

```{r}
library(magrittr) # import all the functions in this package
```

```{r}
lubridate::ymd(....)
```

```{r}
df <-
data.frame(
  types = c('C','C','C','A','C','A','A','B','B','B','B','C'),
  response = c(83,59,54,68,64,88,72,73,66,94,53,55)
)

df$types |> class()

df$types %<>% factor()

df$types |> class()

```

```{r}
df$type |> #read use df$type, to do factor
  factor() |> #to do the table
  table()
```

```{r}
#df$types |> print()
df$type %T>%
  print() |> #read use df$type, to do factor
  factor() |> #to do table
  table()
```

```{r}
c(2, 3, -1) > c(3, 3, 5)
c(2, 3, -1) >= c(3, 3, 5)
```

```{r}
pick <- c(2, 3, -1) >= c(3, 3, 5)
c(2, 3, -1)[pick]
```

```{r}
income <- factor(
  c('20,001-30,000','20,001-30,000','less than 10,000','10,001-20,000','10,001-20,000','20,001-30,000','20,001-30,000','20,001-30,000'),
  levels=c("less than 10,000", "10,001-20,000", "20,001-30,000"),
  ordered = T)

income < "10,001-20,000"
income <= "10,001-20,000" 

pick <- income < "10,001-20,000"
income[pick]
which(pick)
income[which(pick)]
```

```{r}
birthdays <- 
  lubridate::ymd(
    c('2017-11-29','2001-11-07','2011-03-30','2014-03-26','2011-04-20','2014-06-11')
  )

# born after 2002
pick <- birthdays > lubridate::ymd("2002-12-31")

birthdays[pick]
```

```{r}
pick <- c(5, 10, -1) > 0
#how many numbers are larger than zero
sum(pick)#how many element values satisfy the condition(>0)
```

```{r}
pick <- c(5, NA, -1) > 0
print(pick)
sum(pick, na.rm =T)
```

```{r}
pick |> 
  na.omit()
sum()
```

```{r}
johnDoe <- list()
johnDoe$source[[1]] <-
"https://www.moj.gov.tw/2204/2771/2773/76135/post"
johnDoe$source[[2]] <-
"https://www.moj.gov.tw/media/16982/55815125389.xls?mediaDL=true"
googlesheet4::read_sheet(
  johnDoe$source[[2]]
) ->johnDOe$data
```

```{r}
johnDoe$data |>view()
```

```{r}
a=5
```

```{r}
a ==5
```

```{r}
studentGradesInputFromTA <- 
  c(88, 52, 73)
studentGradesFromTeacherCalculation <-
  c(88, 51, 72)
```

```{r}
studentGradesInputFromTA == 
  studentGradesFromTeacherCalculation
```

```{r}
whichIsTheSame <- 
  which(
   pick
  )
print(whichIsTheSame)
```

```{r}
!A ==B #negate A then compare it with B
!(A==B) #compare A with B , then negate the result
```

```{r}
print(studentGradesInputFromTA == 
  studentGradesFromTeacherCalculation)
print(!(studentGradesInputFromTA == 
  studentGradesFromTeacherCalculation))

whichIsDifferent <- 
  which(
    !(studentGradesInputFromTA == 
  studentGradesFromTeacherCalculation)
  )

print(whichIsDifferent)
```

```{r}
!(studentGradesInputFromTA == 
  studentGradesFromTeacherCalculation) # the same as
studentGradesInputFromTA != 
  studentGradesFromTeacherCalculation
```

```{r}
# compare two character vectors
c("A", "B", "C") == c("B", "A", "C")
!(c("A", "B", "C") == c("B", "A", "C"))
c("A", "B", "C") != c("B", "A", "C")

# list is not atomic vector
list("John", 182, 35) == list("John", 182, 40)
```

```{r}
studentGradesInputFromTA2 <- c(
  82, NA, 73
)
studentGradesInputFromTA2 ==
  studentGradesFromTeacherCalculation
```

```{r}
studentGradesFromTeacherCalculation2 <- 
  c(82, NA, 73)

studentGradesInputFromTA2 == 
  studentGradesFromTeacherCalculation2

sum(pick, na.rm=T)

```

```{r}
identical(
  studentGradesInputFromTA2, 
  studentGradesFromTeacherCalculation2)
```

```{r}
studentGradesInputFromTA2
is.numeric(studentGradesInputFromTA2)
is.na(studentGradesInputFromTA2)
```

```{r}
anyNA(studentGradesInputFromTA2)
```

```{r}
bitthdays > lubridate::ymd("2002-12-31")
```

```{r}
birthdays >c(lubridate::ymd("2002-12-31"),
  lubridate::ymd("2002-12-31"), 
  lubridate::ymd("2002-12-31"), 
  lubridate::ymd("2002-12-31"), 
  lubridate::ymd("2002-12-31"), 
  lubridate::ymd("2002-12-31"))

birthdays > rep(lubridate::ymd("2002-12-31"), 6)
```

```{r}
# long vector
long_vctr = c("A", "B", "C", "D")

# short vector
short_vctr = c(1, 2)
# recycle short_vctr
short_vctr_recycles = 
  c(c(1, 2), c(1, 2)) #[1:4]
short_vctr_recycles

# short vector
short_vctr2 = c(1, 2, 3)
# recycle short_vctr
short_vctr_recycles2 = 
  c(c(1, 2, 3), c(1, 2, 3))[1:4]
short_vctr_recycles2
```

```{r}
vc1 =c("Apr.", "May.")
vc2 =c("1", "1")
paste(vc1,vc2)
paste(c("Apr.", "May."), c("1", "1"))
# use recycling
paste(c("Apr.", "May."), "1")
```

```{r}
rep(1998:2021, 2)
.years <- as.character(rep(1998:2021, each=2))
length(.years)

.halfYearDays <-c("01-01", "07-01")

paste(.years, .halfYearDays, sep="-")

```

```{r}
LHS %in% RHS
```

```{r}
# 10 students' majors
majors <- c('economics','law','economics','sociology','sociology','sociology','sociology','economics','statistics','law')
# ? who are from the school of social science ?

# define a set of values that the school contain
set_schoolSocialScience = c("economics", "sociology", "social work")
pickSchoolSocialScience <-
  majors %in% set_schoolSocialScience
print(pickSchoolSocialScience)

which(pickSchoolSocialScience)
```

```{r}
"1995" != "1995 " # even space makes a difference
"economics" != "Economics" # case matters
"台灣" != "臺灣"
```

##4.6.2
```{r}
set.seed(2020)
A<- sample (50:80, 10, replace=T)
print(A)
pick <- A > 60 # is asking 
whichMetTheCondition <- which(pick)
```

```{r}
A[pick]
A[whichMetTheCondition]
```

```{r}
grades <- c(51,70,79,78,67,74,80)
# mean student grades
```

```{r}
pick <- grades > 70 # or
whichIs70plus <- which(pick)
```

```{r}
grades[pick] 
```
*[pick]: retrieve those grades>70
*grades: retrieve grades from those 
*grades[pick]: retrieve from grades for those whose grades >70

```{r}
grades <- c(51,70,79,78,67,74,80)
# same 7 student's gender
gender <- c('female','female','male','male','female','male','male')
```

```{r}
pick <- grades > 70
gender[pick]
```

```{r}
gender[pick] #means
```

*retrieve from from **gender**
*for those whose **grades > 70**

```{r}
dataSet1 <- 
  data.frame(
    school_id = c("001", "002", "003", "004"),
    course1_grade=c(55, 83, 92, 73),
    course2_grade=c(50, 88, 72, 77)
  )
```

For those who have a grade >=80 in course1, what are their grades in course2?
```{r}
# For those who have a grade >=80 in course1
pick <- c(55, 83, 92, 73) >=80
pick <- dataSet1$course1_grade >=80
dataSet1$course2_grade[pick]
#what are their grades in course2
#retrieve their grades in course2
c(50, 88, 72, 77)[pick]
dataSet1$course2_grade[pick]
```

```{r}
pick <- dataSet1$couse1_grade >=80
dataSet1$course2_grade[pick]
```

* For those whose **course1 grades** are >=80,
```{r}
dataSet1$course2_grade[pick] #[pick] conditional retrieval
```

*their **course2 grades** are

***
For those who have course1 grade >=80, what are their records
*A record is a collect of features from the same observation(student)
```{r}
dataSet1 |>names() #|> only R 4.10
names(dataSet1)

pick <- dataSet1$course1_grade >=80
```

```{r}
dataSet1$school_id[pick]
dataSet1$course1_grade[pick]
dataSet1$course2_grade[pick]
```

```{r}
class(dataSet1)
```

* `[.row, .col]`
```{r}
View(dataSet1)
```

```{r}
dataSet1[pick, ]->subsample
View(subsample)
```

*For those whose course1 grades >=80 (`[pick, ]),
*their records (`dataSet1`) are `dataSet1[pick, ]`
**conditional retrieval 
*For those whose..., 
*retrieve xxx.

````{r}
xxx[...]
```

```{r}
subset(dataSet1, course1_grade >=80) ->subsample2
View(subsample2)
#same as
pick <- dataSet1$course1_grade >=80
dataSet1[pick,]
```

```{r}
pick <- c(F, T, NA, T)

c(1, 2, 3, 4)[pick]
```

#4.6.3
```{r}
gender = c('Female','Male','Male', 'Female')
age = c(28,41,42,33)
residence = c('South','South','North', 'North')
color=c("yellow", "pink", "blue", "green")
```

*For those who ...

   *is male
```{r}
pick <- gender =="Male"
```
*For those who ...

   1.are male; AND/OR
   
   2.lives in the south
```{r}
#1.
pick_male <- gender == "Male"
#not 1
!pick_male
#2.
pick_south <- residence =="South"
```

```{r}
# 1. AND 2.
pick_male &pick_south
#1 OR 2
pick_male | pick_south
```

***

```{r}
pick <- list()
```

```{r}
# For those who are "Male": 
pick$male <- gender == "Male"
print(pick$male)
gender
# For those who live in the "South:
pick$south <- residence == "South"
print(pick$south)
residence
```

```{r}
# For those who are "Male" 
#   AND
#   lives in the "South",
pick$male_south <- 
  pick$male & pick$south

print(pick$male_south)

# what are their ages?
age[pick$male_south]
```

For those who are “Male” OR lives in the “South”, what are their ages?
(`[pick$male_south]`), what are their ages `age...`?
***
For those who are “Male” OR lives in the “South”
```{r}
pick$male0south <-
  pick_male | pick$south
```
what are their ages?
```{r}
gender
residence
age
age[pick$male0south]
```

For those who are male or from south, but not both
*exclude `pick$male & pick$south`

```{r}
#pick$male xor pick$south
xor(pick$male, pick$south) ->pick$maleXOsouth

gender
residence
age
age[pick$maleXOsouth]
```

```{r}
pick$male
gender
color
color[pick$male]
color[pick$south]
color[pick$male &pick$south]
color[pick$male | pick$south]
color[xor(pick$male, pick$south)]
```

```{r}
# For those who are male, but not from South
pick$maleXsouth <- 
  pick$male & !pick$south
color[pick$maleXsouth]

# For those who are neither male nor from south
pick$XmaleXsouth <-
  !pick$male & !pick$south
color[pick$XmaleXsouth]
```

```{r}
!(! pick$male) # the same as
pick$male
```

```{r}
!(pick$male & pick$south) # the same as
!pick$male | !pick$south
```

```{r}
!(pick$male | pick$south) # the same as
!pick$male & !pick$south
```

```{r}
any(pick$male)
all(pick$male)
```

```{r}
pick2 <- c(T, T, NA)
any(pick2) # T
all(pick2) #NA
any(na.omit(pick2))
all(na.omit(pick2))
```

```{r}
pick3 <- c(F, F, NA)
any(pick3) #NA
all(pick3) #F
```
